### 2018-04-23 chapters/02-synth.tex
add chapter 4; small changes

Nevertheless, the state of São Paulo has experienced a drastic reduction in homicides during the last years \citep{camargo2007}. The decline is so remarkable that some authors have called it `the great homicide drop' \citep{goertzel2009}. The city of São Paulo, which is currently home to about 11 million inhabitants, provides a telling example. Over a span of only seven years (2000--2007), the number of annual violent deaths in the capital fell from 5,979 to 1,311, a 78\% decrease.\footnote{The homicide statistics cited in this paragraph come from the Centre for the Study of Violence, a research group of the University of São Paulo. Their dataset can be found at the following electronic address: \url{http://nevusp.org/downloads/bancodedados/homicidios/distritossp/num-homicidios-distritos-2000-2007.htm}. Access: March, 2016.} Significantly, São Paulo city became the safest state capital in Brazil \citep{mapa2011}.

@@

Lastly, I estimate another synthetic control using another approach. Here, I employ a Bayesian structural time-series model to verify the stability of the previous results \citep{brodersen2015inferring}. This inference procedure is similar to that described in section \ref{sec:methods} and it also consists of matching pre-treatment values of the unit of interest, São Paulo, to other potential control states. However, in this model only the time trends of the dependent variable are matched. In a sense, this is closer to a traditional differences-in-differences approach, but without the restrictive assumption that the treated and the control cases would follow parallel trends over time \citep[494]{abadie2010}.

@@

The \texttt{R} code below replicates all statistical analyses and graphs included in this chapter. The original data files as well as the final dataset are available at \url{https://github.com/danilofreire/homicides-sp-synth}.

@@

# Prepare dataset
df$state <- as.character(df$state) # required by dataprep()

>>>>

Nevertheless, the state of São Paulo has experienced a drastic reduction in homicides during the last years \citep{camargo2007}. The decline is so remarkable that some authors have called it `the great homicide drop' \citep{goertzel2009}. The city of São Paulo, which is currently home to about 11 million inhabitants, provides a telling example. Over a span of only seven years (2000--2007), the number of annual violent deaths in the capital fell from 5,979 to 1,311, a 78\% decrease.\footnote{The homicide statistics cited in this paragraph come from the Centre for the Study of Violence, a research group of the University of São Paulo. Their data set can be found at the following electronic address: \url{http://nevusp.org/downloads/bancodedados/homicidios/distritossp/num-homicidios-distritos-2000-2007.htm}. Access: March, 2016.} Significantly, São Paulo city became the safest state capital in Brazil \citep{mapa2011}.

@@

Lastly, I estimate another synthetic control using a different approach. I employ a Bayesian structural time-series model to verify the stability of the previous results \citep{brodersen2015inferring}. This inference procedure is similar to that described in section \ref{sec:methods} and it also consists of matching pre-treatment values of the unit of interest, São Paulo, to other potential control states. However, in this model only the time trends of the dependent variable are matched. In a sense, this is closer to a traditional differences-in-differences approach, but without the restrictive assumption that the treated and the control cases would follow parallel trends over time \citep[494]{abadie2010}.

@@

The \texttt{R} code below replicates all statistical analyses and graphs included in this chapter. The original data files as well as the final data set are available at \url{https://github.com/danilofreire/homicides-sp-synth}.

@@

# Prepare data set
df$state <- as.character(df$state) # required by dataprep()

### 2018-04-23 chapters/03-bicho.tex
add chapter 4; small changes

The remainder of this article proceeds as follows. Section \ref{sec:organisation} presents a brief historical overview of the \emph{jogo do bicho}. It describes the necessary conditions for the emergence of the game and presents its basic organisational structure. Section \ref{sec:governance} details the \emph{jogo do bicho}'s governance mechanisms, particularly the strategies employed by vendors to increase trust in markets that operate at the margins of the law. Section \ref{sec:capture} discusses the links between illegal gambling markets, samba schools and the Brazilian state. Section \ref{sec:conclusion} offers some concluding remarks.

@@

\section{Conclusion}
\label{sec:conclusion}

>>>>

The remainder of this article proceeds as follows. Section \ref{sec:institution} presents a brief historical overview of the \emph{jogo do bicho}. It describes the necessary conditions for the emergence of the game and presents its basic organisational structure. Section \ref{sec:governance} details the \emph{jogo do bicho}'s governance mechanisms, particularly the strategies employed by vendors to increase trust in markets that operate at the margins of the law. Section \ref{sec:capture} discusses the links between illegal gambling markets, samba schools and the Brazilian state. Section \ref{sec:conclusion3} offers some concluding remarks.

@@

\section{Conclusion}
\label{sec:conclusion3}

### 2018-04-23 dissertation.tex
add chapter 4; small changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OTHER FORMATTING/LAYOUT DECLARATIONS
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{amssymb,amsmath}
\usepackage{lmodern}
\usepackage{libertine}
\usepackage[libertine]{newtxmath}
\usepackage{inconsolata}
\usepackage[stable]{footmisc}
\usepackage[dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.55}
\usepackage{setspace}
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
\usepackage[backref,pagebackref]{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\cleanlookdateon
\exhyphenpenalty=1000
\hyphenpenalty=1000
\widowpenalty=10000
\clubpenalty=10000
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DECLARATIONS
\chapter*{Declaration}
\noindent 
I hereby declare that except where specific reference is made to the work of others, the contents of this dissertation are original and have not been submitted in whole or in part for consideration for any other degree or qualification in this, or any other university. This dissertation is my own work and contains nothing which is the outcome of work done in collaboration with others, except as specified in the text and Acknowledgements.

@@

% SUMMARY HERE (300 word limit for most subjects):
This dissertation addresses four questions on political and criminal violence. The first essay is an empirical evaluation of a broad set of homicide reduction policies implemented in the state of São Paulo, Brazil. I employ the synthetic control method, a generalisation of differences-in-differences, to compare these measures against an artificial São Paulo. The results indicate a large drop in homicide rates in actual São Paulo when contrasted with the synthetic counterfactual, with about 20,000 lives saved during the period.

@@

In the third essay, co-authored with Gary Uzonyi, we employ extreme bounds analysis and two machine learning methods -- distributed random forests and gradient boosting machines -- to identify the key determinants of state-sponsored violence. We test the sensitivity of 35 variables on a sample of 177 countries from 1945 to 2013. Our results show that GDP per capita and the post-Cold War period are negatively associated with mass killings, whereas previous political turmoils increase the risk of state-led violence. Years since the last episode of mass violence, GDP per capita, urban population, openness to trade, and the number of military personnel make the greatest contribution to our models' out-of-sample predictive power.

@@

% ACKNOWLEDGEMENTS HERE:
I would like to extend thanks to the many people, in many countries, who so generously contributed to the work presented in this dissertation. First and foremost, I want to thank my supervisor, Professor David Skarbek. It was a great privilege and honour to work and study under his guidance, and he has taught me more than I could ever give him credit for here. David was not only an excellent supervisor but also a great friend; his empathy, enthusiasm, and sense of humour made my stay at King's and in London a very pleasant one. I simply could not have wished for a better mentor for my PhD studies. Thank you very much.

@@

The Department of Political Economy at King's College London has provided me with a very stimulating environment in what concerns the extraordinary quality of its academic staff, and that experience will leave marks beyond this thesis. I express my thanks to Professor Florian Foos, Professor Ruben Rufino and all my PhD colleagues for the inspirational discussions and suggestions. Professors Kostas Matakos, John Meadowcroft, Mark Pennington, and Emily Skarbek have supported me in so many ways that I cannot even count. Thank you very much for the opportunities you provided, for your time and skills, and for having faith in me.

I have been very fortunate to work with bright and motivated co-authors during this and other related projects. I am grateful to Rodolpho Bernabel, Gabriel Cepaluni, Diogo Costa, Malte Hendrickx, Robert McDonnell, Umberto Mignozzetti, and Gary Uzonyi for having taught me so much about both scientific research and life in general.

One of the best experiences I have had in the previous years was to become affiliated with the Mercatus Center and the Institute for Humane Studies. There, I have met a community of smart, hard-working and innovative scholars that have in many ways contributed to this research. I would like to express my gratitude to Paul Dragos Aligica, Nigel Ashford, Peter Boettke, Don Boudreaux, Brandon Brice, John Buchmann, Jeff Carroll, Chris Coyne, Courtney Dunn, Bobbi Herzberg, Marcus Hunt, Nick Jacobs, Isaac Jilbert, Peter Lipsey, Ian Madison, Kelly Nelson, Gabriel Oliva, Raj Patel, Blaž Remic, Virgil Storr, Alienor van den Bosch, Richard Wagner, and Larry White for all the engaging conversations. I have learned so much from you.

Completing this work would have been all the more difficult were it not for the support and constructive criticism provided by my friends and colleagues. Specially, I would like to thank Veronica Adaeva, André Amaro, Aline Bagatin, Fábio Barros, Michel Hulmann, Guilherme Kerr, Maurício Pantaleão, Letícia Sene, and the participants of academic seminars at King's College London, the Mercatus Center at George Mason University, and the Prometheus Institute in Berlin. I am deeply indebted to them for their encouragement, insightful comments, and hard questions.

I am deeply thankful to several present and past members of the Friedrich Naumann Foundation for Freedom. Many years ago, I was received a scholarship to attend a short course in Germany, and that first experience abroad transformed my life ever since. I then decided to pursue my post-graduate studies in Europe and it was one of the best choices I have made. The present work is, to a large extent, a result of the push I received from the Naumann Foundation almost 10 years ago. I owe a great debt of gratitude to Cidália Achten, Gulmina Bilal, Rainer Erkers, Beate Forbriger, Verena Gierszewski, Birgit Lamm, Monica Lehmann, Stefan Melnik, and Wulf Pabst for their support. As time goes on, it is easy to see the huge impact they have had on my academic career.

A very special word of thanks goes for my mother Rosali and my grandmother Maria. They have been absolutely wonderful over the years and have gone above and beyond to provide me with a good education. This dissertation is dedicated to them. \textit{Nós conseguimos}.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OTHER FORMATTING/LAYOUT DECLARATIONS
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{amssymb,amsmath}
\usepackage{lmodern}
\usepackage{libertine}
\usepackage[libertine]{newtxmath}
\usepackage{inconsolata}
\usepackage[stable]{footmisc}
\usepackage[dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.55}
\usepackage{setspace}
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}
\usepackage[backref,pagebackref]{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}
\usepackage{dcolumn}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\cleanlookdateon
\exhyphenpenalty=1000
\hyphenpenalty=1000
\widowpenalty=10000
\clubpenalty=10000
\usepackage[fencedCode,inlineFootnotes,citations,definitionLists,hashEnumerators,smartEllipses,hybrid]{markdown}
\markdownSetup{rendererPrototypes={
   link = {\href{#2}{#1}},
   image = {\begin{figure}[hbt!]
     \centering
     \includegraphics{#3}%
     \ifx\empty#4\empty\else
     \caption{#4}%
     \fi
     \label{fig:#1}%
     \end{figure}}
}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DECLARATIONS
\chapter*{Declaration}
\noindent 
I hereby declare that except where specific reference is made to the work of others, the contents of this thesis are original and have not been submitted in whole or in part for consideration for any other degree or qualification in this, or any other university. This thesis is my own work and contains nothing which is the outcome of work done in collaboration with others, except as specified in the text and Acknowledgements.

@@

% SUMMARY HERE (300 word limit for most subjects):
This thesis addresses three questions on political and criminal violence. The first essay is an empirical evaluation of a broad set of homicide reduction policies implemented in the state of São Paulo, Brazil. I employ the synthetic control method, a generalisation of differences-in-differences, to compare these measures against an artificial São Paulo. The results indicate a large drop in homicide rates in actual São Paulo when contrasted with the synthetic counterfactual, with about 20,000 lives saved during the period.

@@

In the third essay, co-authored with Gary Uzonyi, we employ extreme bounds analysis and distributed random forests to identify the key determinants of state-sponsored violence. We test the sensitivity of 40 variables on a sample of 177 countries from 1945 to 2013. The results show that GDP per capita, democracy, and the post-Cold War period are negatively associated with mass killings, whereas ethnic diversity, civil wars, and previous political turmoils increase the risk of state-led violence. GDP per capita, years since the last episode of mass violence, urban population, ethnic polarisation, and the number of military personnel make the greatest contribution to the models' out-of-sample predictive power.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% ACKNOWLEDGEMENTS HERE:
I would like to extend thanks to the many people, in many countries, who so generously contributed to the work presented in this thesis. First and foremost, I want to thank my supervisor, Professor David Skarbek. It was a great privilege and honour to work and study under his guidance, and he has taught me more than I could ever give him credit for here. David was not only an excellent supervisor but also a great friend; his empathy, enthusiasm, and sense of humour made my stay at King's College a very pleasant one. I simply could not have wished for a better mentor for my PhD studies. Thank you very much.

@@

The Department of Political Economy at King's College London has provided me with a very stimulating environment in what concerns the extraordinary quality of its students and  academic staff, and that experience will leave marks beyond this thesis. I express my thanks to Professor Florian Foos, Professor Rubén Ruiz-Rufino, and all my PhD colleagues for the inspirational discussions and suggestions. David Muchlinski has taught me many things about machine learning and I am very thankful for that. Professors Kostas Matakos, John Meadowcroft, Mark Pennington, and Emily Skarbek have supported me in so many ways that I cannot even count. Thank you very much for the opportunities you provided, for your time and skills, and for having faith in me.

I have been very fortunate to work with bright and motivated co-authors during this and other related projects. I am grateful to Rodolpho Bernabel, Gabriel Cepaluni, Diogo Costa, Guilherme Jardim Duarte, Malte Hendrickx, Robert McDonnell, Umberto Mignozzetti, and Gary Uzonyi for having taught me so much about both scientific research and life in general.

One of the best experiences I have had in the previous years was to become affiliated with the Mercatus Center and the Institute for Humane Studies. There, I have met a community of smart, hard-working and innovative scholars that have in many ways contributed to this research. I would like to express my gratitude to Paul Dragos Aligica, Nigel Ashford, Peter Boettke, Don Boudreaux, Brandon Brice, John Buchmann, Jeff Carroll, Chris Coyne, Courtney Dunn, Luke Foster, Bobbi Herzberg, Marcus Hunt, Nick Jacobs, Isaac Jilbert, Peter Lipsey, Ian Madison, Kelly Nelson, Gabriel Oliva, Raj Patel, Pablo Prieto, Blaž Remic, Virgil Storr, Alienor van den Bosch, Richard Wagner, and Larry White for all the engaging conversations. I have learned so much from you.

Completing this work would have been all the more difficult were it not for the support and constructive criticism provided by my friends. Specially, I would like to thank Veronica Adaeva, André Amaro, Paulo Roberto Araujo, Aline Bagatin, Fábio Barros, Michel Hulmann, Guilherme Kerr, Maurício Pantaleão, Letícia Sene, and the participants of academic seminars at King's College London, the Mercatus Center at George Mason University, and the Prometheus Institute in Berlin. I am deeply indebted to them for their insightful comments and hard questions.

I am also deeply thankful to several present and past members of the Friedrich Naumann Foundation for Freedom. Some years ago, I had received a scholarship to attend a short course in Germany, and that first experience abroad has transformed my life ever since. The present work is, to a large extent, a result of the push I received from the Naumann Foundation almost 10 years ago. I owe a great debt of gratitude to Cidália Achten, Gulmina Bilal, Rainer Erkers, Beate Forbriger, Verena Gierszewski, Birgit Lamm, Monica Lehmann, Stefan Melnik, and Wulf Pabst for their support. As time goes on, it is easy to see the huge impact they have had on my academic career.

A very special word of thanks goes for my mother Rosali and my grandmother Maria. They have been absolutely wonderful over the years and have gone above and beyond to provide me with a good education. This thesis is dedicated to them. \textit{Nós conseguimos}.

@@

%-----------------------------------------------------
% Chapter 4: Mass Killings
%-----------------------------------------------------

\input{chapters/04-killings}
